Understanding the Lomb-Scargle Periodogram

### 2017-02-03 PracticalLombScargle.tex
Add section on correlated noise

One important caveat of the Lomb-Scargle approach is that the statistical
guarantees only apply when the observations have {\it uncorrelated white
noise}; data with more complicated noise characteristics must be treated
more carefully; see, e.g., \citep{Vio2010}.

@@

@@

\subsubsection{Computing the Window Function}
The window power spectrum can be computed directly from the
delta-function representation of the window function;
From \eq{FT-def}, \eq{power-spectrum}, and \eq{nonuniform-window},
we can write
\begin{equation}
  \mathcal{F}\{W_{\{t_n\}}\}(f) = \sum_{n=1}^{N} e^{-2\pi i f t_n}
  \eqlabel{nonuniform-window-FT}
\end{equation}
For uniformly-spaced observing times $t_n$, the symmetry leads to cancellations
so that the window transform is a clean Dirac comb.
For non-uniform observing times, this symmetry is broken and the window
transform will be much more ``nosiy''.
for observation times $\{t_n\}$ is
\begin{equation}
  \mathcal{P}_W(f;\{t_n\}) = \left|\sum_{n=1}^{N} e^{-2\pi i f t_n}\right|^2
  \eqlabel{window-power-analytic}
\end{equation}
Comparing this to \eq{schuster-periodogram}, we see that this is essentially
the Schuster periodogram for data $g_n=1$ at all times $t_n$;
with that in mind one quick way to compute the window power is to use
a standard Lomb-Scargle periodogram on a series of unit measurements,
making sure to {\it not} pre-center the data or to use a floating mean model.
This method is how window power spectra have been computed throughout this
paper.

\subsubsection{Deconvolution and CLEANing}
With this ability to compute the window function, we might hope to be able to
use it {\it quantitatively} to remove spurious peaks from the observed
power spectrum
Recall from \eq{g-nonuniform} that the observed data is a point-wise product
of the underlying function and the survey window:
\begin{equation}
  g_{obs}(t) = g_{true}(t)W_{\{t_n\}}(t),
\end{equation}
and the convolution theorem tells us that the observed Fourier transform is
a convolution of the true transform and the window transform:
\begin{equation}
  \mathcal{F}\{g_{obs}\} = \mathcal{F}\{g_{true}\} \ast \mathcal{F}\{W_{\{t_n\}}\}
\end{equation}
Given this relationship, we might hope to be able to invert this
convolution to recover $\mathcal{F}\{g_{true}\}$ directly.
For example, we could write:
\begin{eqnarray}
  \mathcal{F}\{g_{true}\} &=& \mathcal{F}\{g_{obs}/W_{\{t_n\}}\}\nonumber\\
                   &=& \mathcal{F}\{g_{obs}\} \ast \mathcal{F}\{1/W_{\{t_n\}}\}
\end{eqnarray}
Because of the localization of observations, $W_{\{t_n\}}(t)$ is zero for most
values of $t$ and so $1/W_{\{t_n\}}(t)$ and its Fourier transform are not well
defined. For this reason, direct deconvolution is not an option in most
cases of interest---in other words, the deconvolution problem is ill-posed
and does not have a unique solution.

>>>>

One important caveat of the Lomb-Scargle approach is that the statistical
guarantees only apply when the observations have {\it uncorrelated white
noise}; data with more complicated noise characteristics must be treated
more carefully; see, e.g., \citep{Vio2010} or the Least Squares approach
discussed in \sect{extensions-observational-noise}.

@@

This also suggests a convenient way to construct a periodogram that accounts
for correlated noise across observations; denoting the  $N\times N$ noise
covariance matrix for $N$ observations by $\Sigma$, and constructing the
vectors $\vec{y} = [y_1, y_2,\cdots y_n]^T$
and $\vec{y}_{model} = [y_{model}(t_1),y_{model}(t_1),\cdots y_{model}(t_n)]^T$,
then the $\chi^2$ statistic can be written
\begin{equation}
  \chi^2(f) = (\vec{y}-\vec{y}_{model})^T\Sigma^{-1}(\vec{y}-\vec{y}_{model})
  \eqlabel{chi2-with-correlated-errors}
\end{equation}
In the case of uncorrelated errors, the off-diagonal terms of $\Sigma$
are zero, and this expression reduces to \eq{chi2-with-errors}.
This is similar in spirit to the approach to correlated noise developed
by \citet{Vio2010} from the Fourier perspective, and is in fact identical
in the case of the ``zero-mean colored noise'' example considered therein.

@@

\subsubsection{Computing the Window Function}
The window power spectrum can be computed directly from the
delta-function representation of the window function;
From \eq{FT-def}, \eq{power-spectrum}, and \eq{nonuniform-window},
we can write
\begin{equation}
  \mathcal{F}\{W_{\{t_n\}}\}(f) = \sum_{n=1}^{N} e^{-2\pi i f t_n}
  \eqlabel{nonuniform-window-FT}
\end{equation}
For uniformly-spaced observing times $t_n$, the symmetry leads to cancellations
so that the window transform is a clean Dirac comb.
For non-uniform observing times, this symmetry is broken and the window
transform will be much more ``nosiy''.
for observation times $\{t_n\}$ is
\begin{equation}
  \mathcal{P}_W(f;\{t_n\}) = \left|\sum_{n=1}^{N} e^{-2\pi i f t_n}\right|^2
  \eqlabel{window-power-analytic}
\end{equation}
Comparing this to \eq{schuster-periodogram}, we see that this is essentially
the Schuster periodogram for data $g_n=1$ at all times $t_n$;
with that in mind one quick way to compute a qualitatively correct window
power is to use a standard Lomb-Scargle periodogram on a series of unit
measurements, making sure to {\it not} pre-center the data or to use a
floating mean model.
As \citep{Scargle82} notes, this is {\it not} the true window---it differs
from the true window just as the Lomb-Scargle periodogram differs from the
Schuster periodogram---but in practice it is accurate enough to give
useful insight into the window function's important features.
This method is how window power spectra has been computed throughout this
paper.

\subsubsection{Deconvolution and CLEANing}
With this ability to compute the window function, we might hope to be able to
use it {\it quantitatively} to remove spurious peaks from the observed
power spectrum
Recall from \eq{g-nonuniform} that the observed data is a point-wise product
of the underlying function and the survey window:
\begin{equation}
  g_{obs}(t) = g(t)W_{\{t_n\}}(t),
\end{equation}
and the convolution theorem tells us that the observed Fourier transform is
a convolution of the true transform and the window transform:
\begin{equation}
  \mathcal{F}\{g_{obs}\} = \mathcal{F}\{g\} \ast \mathcal{F}\{W_{\{t_n\}}\}
\end{equation}
Given this relationship, we might hope to be able to invert this
convolution to recover $\mathcal{F}\{g\}$ directly.
For example, we could write:
\begin{eqnarray}
  \mathcal{F}\{g\} &=& \mathcal{F}\{g_{obs}/W_{\{t_n\}}\}\nonumber\\
                   &=& \mathcal{F}\{g_{obs}\} \ast \mathcal{F}\{1/W_{\{t_n\}}\}
\end{eqnarray}
Because of the localization of observations, $W_{\{t_n\}}(t)$ is zero for most
values of $t$ and so $1/W_{\{t_n\}}(t)$ and its Fourier transform are not well
defined. For this reason, direct deconvolution is not an option in most
cases of interest---in other words, the deconvolution problem is ill-posed
and does not have a unique solution.
